Manuscript Figures

### 2021-03-17 src/manuscript_figures.Rmd
add survival benchmark, start luad_vs_lusc analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE)
# todo: pass this as argument
settings <- yaml::read_yaml('/data/local/buyar/arcas/pancancer_multiomics_manuscript/settings.yaml')
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
theme_set(ggpubr::theme_pubclean())

# todo: move paths to settings file
inputDir <- '/data/local/buyar/arcas/pancancer_multiomics_manuscript/output'
geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'
FigureFolder <- '/data/bimsbstatic/public/akalin/buyar/manuscript_figures_arcas'
assay_dir <- settings$assay_output$folder
surv <- do.call(rbind, readRDS(settings$surv))
projectDict <- sapply(yaml::read_yaml(settings$projectDict), 
                      function(x) gsub(' ', '', unlist(strsplit(x, ','))))
tcga_subtypes <- process_tcga_subtypes()

Figures <- list() # accummulate all figures in this
SuppFigures <- list() # accumulate all supplementary figures in this
Legends <- list() # figure legend text
TOOLS <- c('pca_factors', 'maui_factors', 'mofa_factors', 'mcia_factors')

@@

Fig <- plot_grid(plot_grid(p1, p2, labels = 'AUTO', ncol = 1),
                  p3, ncol = 2, rel_widths = c(2,3), labels = c('','C'))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.cluster_stability_metrics.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name),        
       plot = Fig, units = 'in', 
       width = 10, height = 8)
Legends[[fig_count]] <- "Comparison of unsupervised clustering performances based on cluster homogeneity (Silhouette) in A) and cluster stability (Average Proportion of Non-Overlap (APN) in B). Each point represents a clustering score for 132 k-means clustering results (21 cancer types x for 6 different values of k from 3 to 8) based on the top 10 LFs/PCs derived from each tool. C) Same as in panels A and B), but split by the value of k used in clustering."
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE, message = FALSE)
# todo: pass this as argument
settings <- yaml::read_yaml('/data/local/buyar/arcas/pancancer_multiomics_manuscript/settings.yaml')
source(settings$utility_script)
library(data.table)
library(ggplot2)
library(pbapply)
library(ggrepel)
library(ggpubr)
library(cowplot)
theme_set(ggpubr::theme_pubclean())
geneSetScoresFolder <- '/data/local/buyar/arcas/subtyping_paper/data/geneset_scores/singscore'
FigureFolder <- '/data/bimsbstatic/public/akalin/buyar/manuscript_figures_arcas'

inputDir <- settings$pipeline_output
assay_dir <- settings$assay_output$folder
surv <- do.call(rbind, readRDS(settings$surv))
projectDict <- sapply(yaml::read_yaml(settings$projectDict), 
                      function(x) gsub(' ', '', unlist(strsplit(x, ','))))
tcga_subtypes <- process_tcga_subtypes()

TOOLS <- c('pca_factors', 'maui_factors', 'mofa_factors', 'mcia_factors')

Figures <- list() # accummulate all figures in this
SuppFigures <- list() # accumulate all supplementary figures in this
Legends <- list() # figure legend text

@@

# survival prediction accuracy benchmarking
# find survival-predictive LFs by univariate cox models and use them in RSF
cl <- parallel::makeCluster(20)
parallel::clusterExport(cl, varlist = c('LFs', 'experiments', 'settings', 'surv'))
surv.stats <- do.call(rbind, sapply(simplify = F, names(LFs), function(tool) {
  message("survival benchmark for ",tool)
  do.call(rbind, pbapply::pbsapply(cl = cl, simplify = F, experiments, function(experiment) {
    message(experiment)
    source(settings$utility_script)
    require(data.table)
    require(survminer)
    M <- LFs[[tool]][[experiment]]
    if(tool == 'maui_factors') {
      M <- remove_redundant_variables(M, 99)
    }
    dt <- data.table('cindex.pfi' = NA, 'cindex.os' = NA, 'experiment' = experiment, 'method' = tool)
    # compute pfi stats
    vars <- get_prognostic_covariates(surv, clin = NULL, explored_covariates = M,
                                      event_col = 'PFI', duration_col = 'PFI.time')
    if(!is.null(vars)) {
      v <- head(vars[order(pval)][pval < 0.05][['variable']], 10)
      fit <- fit_RSF(surv, M[,v], 1000, 'PFI.time', 'PFI')
      dt$cindex.pfi <- round(1 - mean(fit$err.rate, na.rm = T), 2)
    }
    vars <- get_prognostic_covariates(surv, clin = NULL, explored_covariates = M,
                                      event_col = 'OS', duration_col = 'OS.time')
    if(!is.null(vars)) {
      v <- head(vars[order(pval)][pval < 0.05][['variable']], 10)
      fit <- fit_RSF(surv, M[,v], 1000, 'OS.time', 'OS')
      dt$cindex.os <- round(1 - mean(fit$err.rate, na.rm = T), 2)
    }
    return(dt)
  }))
}))
parallel::stopCluster(cl)

dt <- melt(surv.stats, measure.vars = c('cindex.pfi', 'cindex.os'))
dt[is.na(value)]$value <- min(dt$value, na.rm = T) # 
dt$method <- toupper(gsub("_factors", "", dt$method))
dt$project <- unlist(sapply(strsplit(dt$experiment, "\\."), function(x) x[1]))
dt$endpoint <- ifelse(dt$variable == 'cindex.pfi', 'Progression Free Survival', 
                      'Overall Survival')
# PCA/MOFA win, compare against PCA
comps <- lapply(setdiff(unique(dt$method), 'PCA'), function(x) c('PCA', x))
p4 <- ggboxplot(dt, x = 'method', y = 'value', color = 'method',
                add = 'jitter', facet.by = 'endpoint') +
  stat_compare_means(method = 'wilcox.test', paired = TRUE,
                     comparisons = comps) +
  labs(y = "Harrel's C-index", x = '') +
  theme(legend.position = 'none')

Fig <- plot_grid(plot_grid(p1, p2, labels = c('A', 'B'), ncol = 1),
                 plot_grid(p3, p4, labels = c('C', 'D'), ncol = 1), 
                 ncol = 2, rel_widths = c(2,3))

fig_count <- fig_count + 1
Figures[[fig_count]] <- Fig
fig.name <- paste0('Figure', fig_count, '.cluster_stability_metrics.pdf')
names(Figures)[fig_count] <- fig.name
ggsave(file.path(FigureFolder, fig.name),        
       plot = Fig, units = 'in', 
       width = 10, height = 10)
Legends[[fig_count]] <- "Comparison of unsupervised clustering performances based on cluster homogeneity (Silhouette) in A) and cluster stability (Average Proportion of Non-Overlap (APN) in B). Each point represents a clustering score for 132 k-means clustering results (21 cancer types x for 6 different values of k from 3 to 8) based on the top 10 LFs/PCs derived from each tool. C) Same as in panels A and B), but split by the value of k used in clustering. D) Survival outcome prediction accuracy for 21 cancer types by each tool's top 10 survival-related factors based on Overall Survival or Progression Free Survival endpoints."
```

@@

```{r}
# Analysis of non-small-cell lung cancer types (LUSC vs LUAD)
# reference study: https://www.nature.com/articles/s41598-020-77284-8
M <- remove_redundant_variables(LFs$maui_factors$nsclc.cnv_gex_meth_mut, perc = 99)
ct <- surv[match(rownames(M), bcr_patient_barcode)]$project

# find type-specific LFs
dt <- get_factor_specific_variables(M, ct)
top_vars <- get_top_by_label(dt[padj < 0.05][order(padj)], 'ref_cl', 10)
s <- unique(top_vars$variable)
pheatmap(M[,s], show_rownames = F)
plot_tsne(M[,s], ct)
plot_umap(M, ct)
plot_tsne(M, ct)
clust <- get_spectral_clusters(M, k_vals = 2:6)
clin <- process_clinical_covariates(rownames(M))
prog_vars <- get_prognostic_covariates(surv, clin, M)
prog.clust <- get_kmeans_subtypes(M[,unique(prog_vars$variable)], k_vals = 2:8, min_cluster_size = 0)

which.min(sapply(colnames(prog.clust), function(x) {
  plot_survival(surv, prog.clust[,x,drop=F], return_pval = T)
}))

plot_survival(surv, prog.clust[,'kmeans_3', drop = F])
plot_survival(surv, data.frame('ct' = ct, row.names = rownames(M)), 
              'PFI.time', 'PFI')

df <- plot_tsne(M[,unique(prog_vars$variable)], ct, returnData = T)
df$pfi_subtype <- as.factor(prog.clust[rownames(df),'kmeans_3'])

ggplot(df, aes(x = tSNE1, y = tSNE2)) + 
  geom_point(aes(color = as.factor(pfi_subtype))) +
  facet_wrap(~ group, nrow = 1 )

ggplot(df, aes(x = tSNE1, y = tSNE2)) + 
  geom_point(aes(color = group)) +
  facet_wrap(~ as.factor(pfi_subtype), nrow = 1 )

scores <- sapply(simplify = F, c('msigdb_hallmarks', 'msigdb_canonical_pathways', 'cancersea'), 
       function(x) {
         scores <- import_geneset_scores(geneSetScoresFolder, x)
         S <- scores[intersect(rownames(M), rownames(scores)),]
         return(S)
       })

pathways.prog <- get_factor_specific_variables(scores$cancersea, df$pfi_subtype)[padj < 0.05][order(padj)]
pathways.ct <- get_factor_specific_variables(scores$cancersea, df$group)[padj < 0.05][order(padj)]

get_top_by_label(pathways.ct[padj < 0.05][order(padj)], 'ref_cl', 10)

```
